Hard Statistics in Economics: Investigating Economic Drivers in Individual Confidence in the DOW in the US

### 2024-12-04 paper/paper.qmd
two cross-references not working

All data analysis was done through R [@citeR] with the aid of the following packages: `tidyverse` [@citetidyverse], `dplyr` [@dplyr], `ggplot2` [@citeggplot2], `knitr` [@citeknitr], `vars` [@vars], `arrow` [@arrow], `kableExtra` [@citekableExtra], and `janitor` [@citejanitor]. Moreover, the FRED API was accessed using the package `fredr` [@fredr] and the shiny web app associated with this paper was made with the aid of the `shiny` package [@citeshiny].

@@

## Economic Variables Are Poor Predictors of Market Confidence
Despite spending time to choose a model with normalized errors and the smallest mean square errors based on predictions, this model still does not perform well and it's interpretation lacks good understanding from an investors standpoint. It indicates that these economic factors are not at all good predictors of market confidence. It shows that the market is not influenced by these economic factors alone and there is a much more complex and intricate system of factors that play into how an investor may feel about the future. It would have been nice to find that the most common economic statistics, like GDP growth or the VIX, play a large role in ones confidence in the market, but there is no obvious linear relationship.

@@

The second step in creating the model was to determine the lag in our model. That is determine the number of lags for each variable we wanted to include. To do this we did employed an `acf.test` fromthe `vars` package which tests for the optimal lag in each variable. It turns out that the optimal lag among all variables was 9. From there we created two models, one with lag 1 and the other with lag 9. That means the second model included a $t-n$ term on each coefficient for $n={1,...,9}$. We found that they both had a mean square error of approximately 7.33.

>>>>

All data analysis was done through R [@citeR] with the aid of the following packages: `tidyverse` [@citetidyverse], `dplyr` [@dplyr], `ggplot2` [@citeggplot2], `knitr` [@citeknitr], `vars` [@vars], `arrow` [@arrow], `kableExtra` [@citekableExtra], and `testthat` [@testthat]. Moreover, the FRED API was accessed using the package `fredr` [@fredr].

@@

## Economic Variables Are Poor Predictors of Market Confidence
Despite spending time to choose a model with normalized errors and the smallest mean square errors based on predictions, this model still does not perform well and it's interpretation lacks good understanding from an investors standpoint. As discussed in the data section, we do not expect this model to perform perfectly given that is it attempting to predict one-year confidence indexes during the peak of COVID-19, but nonetheless we cannot conclude it has great predictive performance. Its performance indicates that these economic factors are not all good predictors of market confidence. It shows that the market is not influenced by these economic factors alone and there is a much more complex and intricate system of factors that play into how an investor may feel about the future. It would have been nice to find that the most common economic statistics, like GDP growth or the VIX, play a large role in ones confidence in the market, but there is no obvious linear relationship.

@@

The second step in creating the model was to determine the lag in our model. That is determine the number of lags for each variable we wanted to include. To do this we did employed an `acf.test` from the `vars` package which tests for the optimal lag in each variable. It turns out that the optimal lag among all variables was 9. From there we created two models, one with lag 1 and the other with lag 9. That means the second model included a $t-n$ term on each coefficient for $n={1,...,9}$. We found that they both had a mean square error of approximately 7.33.
